Numerical simulation of the Cordilleran ice sheet through the last glacial cycle

### 2014-05-21 cordillera-cycle.tex
Update multi-record snapshots.
Add intro first shot.

The Cordilleran ice sheet has a long history of explorations and geological
investigations. However, the complexity of the evidence is such that no ice-
sheet scale reconstructions are yet possible. Here we use numerical modelling,
calibrated with field evidence to attempt such a reconstruction.

@@

All six simulations consistently produce two major glaciations during MIS~4 and
MIS~2 (Figs.~\ref{fig:timeseries} and~\ref{fig:multirec}).
However the timing is much different. Notably the oceanic core forcing result
in much too early last glacial maximum.

@@

% fig:multirec-timeseries
\begin{figure}
  \includegraphics{timeseries}
  \caption{Multi-record timeseries. In construction.
           \todo{Add NGRIP}
           \idea{Use same colors but different patterns for similar records}}
  \label{fig:timeseries}
\end{figure}

% fig:multirec-snapshots
\begin{figure}
  \includegraphics{multirec}
  \caption{Multi-record snapshots. In construction.
           \todo{Add NGRIP}}
  \label{fig:multirec}
\end{figure}

>>>>

During the last glacial cycle, glaciers and ice caps of the North American
Cordillera have been more extensive than today. At the peak of glaciation, a
contiguous blanket of ice extended from Denali Peak in the north to the Puget
Sound in the south. This blanket is known as the former Cordilleran ice sheet.

For more than a century, exploration and geological investigation of the
Cordillera lead to a large collection of evidence of the former ice cover. This
evidence consists of mapped boundaries of the former ice extent, direction of
past sliding velocities and location of former melt-water streams, as well as
carbon dating and cosmogenic dating of glacier retreat.

Field-based evidence allowed to reconstruct, with a high degree of confidence,
the maximal extent attained by the Cordilleran ice sheet during the last
glacial cycle. However, former ice thickness, and in turn the ice sheet's
contribution to the last glacial maximum sea-level low-stand, remains poorly
constrained. Moreover, our understanding of the deglaciation phase is
restricted to a regional level, while little is known about the ice sheet
history preceding the last glacial maximum extent. Although time-evolving,
whole ice-sheet reconstructions of glacial advance and retreat patters are
available for the Laurentide and Eurasian ice sheet, this is not the case for
the Cordilleran ice sheet.

The present study aims to use a numerical ice sheet model calibrated against
field-based evidence to attempt a quantitative reconstruction of the
Cordilleran ice sheet history through the last glacial cycle. Although
numerical modelling has been identified as a tool to improve our understanding
of Cordilleran ice sheet more than twenty years ago, the ubiquitously
mountainous topography of the region has presented a major challenge to its
application, which is only overcome by recent development of numerical ice
sheet models and underlying scientific tools. In addition to the topographic
complexity, strong climatic contrasts characteristic of the North American
Cordillera require the use of high-resolution temperature and precipitation
forcing as input data to the ice sheet model.

Here we use the Parallel Ice Sheet Model (PISM), forced by a high-resolution
climatology from the North American Regional Reanalysis, to model the
Cordilleran ice sheet through the last glacial cycle. Because past climate
is mostly unknown, palaeoclimate forcing is applied in the simplified form of
six time-dependent temperature offset time-series. These time-series were
obtained by scaling six different palaeo-temperature reconstructions from
proxy records around the globe, including two \chem{\delta^{18}O} records from
Greenland ice cores, two \chem{\delta^{18}O} records from Antarctic ice cores,
and two alkenone unsaturation index records from Northwest Pacific oceanic
sediment cores. The output from the model is compared to geomorphological
evidence in terms of timing of glaciation, extent of glaciation and patterns
of deglaciation.

% Greenland Ice Core Project (GRIP)
% North Greenland Ice Core Project (NGRIP)
% European Project for Ice Coring in Antarctica (EPICA)
% Lake Vostok (Vostok)
% Ocean Drilling Program (ODP)

@@

All six simulations consistently produce two major glaciations during MIS~4 and
MIS~2 (Figs.~\ref{fig:timeseries} and~\ref{fig:snapshots}).
However the timing is much different. Notably the oceanic core forcing result
in much too early last glacial maximum.

@@

% fig:timeseries
\begin{figure}
  \includegraphics{timeseries}
  \caption{Multi-record timeseries. In construction.
           \todo{Add NGRIP}
           \idea{Use same colors but different patterns for similar records}}
  \label{fig:timeseries}
\end{figure}

% fig:snapshots
\begin{figure}
  \includegraphics{snapshots}
  \caption{Multi-record snapshots. In construction.
           \todo{Add NGRIP}}
  \label{fig:snapshots}
\end{figure}
